---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# 1. Loading Data

### Data loaded into dataframe

```{r}

rm(list=ls())

coal_df = read.csv(file = 'dataset_Nicholas-Tony.csv')

```
# 2. Mean and Variance of each variable


### Thickness

```{r}
cat("Mean thickness: ", mean(coal_df$Thickness))
cat("\nVariance of thickness: ", var(coal_df$Thickness))
```


### Accumulation

```{r}
cat("Mean accumulation: ", mean(coal_df$Accumulation))
cat("\nVariance of accumulation: ", var(coal_df$Accumulation))
```

Accumulation has a very large variance. In the Gaussian case, this would result in negative values for accumulation. Since accumulation is a physical nonnegative quantity, it cannot be distributed in this way. 
However, the large variance is consistent with the given multimodal sum-of-Gaussians distribution.

### Lignite Quality Index

```{r}
cat("Mean lignite quality: ", mean(coal_df$LQ))
cat("\nVariance of lignite quality: ", var(coal_df$LQ))
```

Only 29% of the observations have quality index 1. 79% have quality index 0. 
Variance matches Bernoulli RV with $p = 0.29$, $p(1-p) = 0.2059$.


# 3. Histograms of each variable

```{r}
hist(coal_df$Thickness, breaks = 10)
```
The curve is bell-shaped with a single mode around the average at 10. The curve shows a slight negative skew. Median is 10.38, slightly higher than mean. 

Given model for Thickness is $\mu = 10$, $\sigma^2 = 4$.

Testing validity of given model with 1-sample Kolmogorov-Smirnov test:

```{r}
ks.test(coal_df$Thickness, pnorm, 10, 2)
```
The p-value of 0.06661 exceeds the significance 0.05, so we do not reject the hypothesis that the sample was drawn from this distribution.

```{r}
hist(coal_df$Accumulation, breaks=10)
```

The accumulation is bimodal, one mode centered at low values with low variance, and another with a high mean and high dispersion. 

```{r}

coal_df[coal_df$LQ == 1,]

```



# 4. Conditional simulation of thickness T on discretization grid


Preparing the RGeostats db object

```{r}
library(RGeostats)

#Changing the quality index to avoid issue with 0 value

if (any(coal_df$LQ == 0)){
  coal_df$LQ = coal_df$LQ + 1
}


coal_db <- db.create(coal_df)

coal_db = db.locate(coal_db,3:4,"x")
coal_db = db.locate(coal_db,5,"z")

```


Simulating the thickness, modeled as a Gaussian random function with parameters $\mu = 10$ and $\sigma^2 = 4$.

```{r warning=FALSE}


sim_thickness <- function(indb, nsim = 1){
  
  indb = db.locate(indb, "Thickness", "z")
  
  Tmodel=model.create(vartype="Gaussian",range=20,sill=4,ndim=2)
  outdb = db.create(x0=c(0,0),dx=c(1,1),nx=c(200,100))
  neighb = neigh.create(ndim = 2, type = 0)
  simu <- simtub(nbtuba = 450, dbin=indb,dbout=outdb,model=Tmodel,neigh = neighb,mean = c(10),seed=933033161, nbsimu = nsim)
  return(simu)

}

simu = sim_thickness(coal_db)

plot(simu, xlab="X", ylab="Y", title="Thickness (Simulated)")
plot(coal_db, add=T)
```


# 5. Conditional simulation of LQ (lignite quality index)

LQ is modelled as a thresholded GRF with constant proportion 0.5. The underlying GRF is standard Gaussian ($\mu = 0$ and $\sigma^{2} = 1$) and covariance is Gaussian with range 50. 

```{r warning=FALSE}


sim_lq <- function(indb, nsim = 1){
  
  indb = db.locate(indb, "LQ", "z")
  props = c(0.5,0.5)
  rule = rule.create(c("S", "F1", "F2"))
  voisin = neigh.create(type=0)
  outdb = db.create(x0=c(0,0),dx=c(1,1),nx=c(200,100))
  gaus = model.create("Gaussian", range=50)
  simlqin = simpgs(dbout=outdb, dbin=indb, rule=rule,props=c(0.5, 0.5), model1=gaus, neigh = voisin,seed=573738494, nbsimu = nsim)
  return(simlqin)

}

simlq0 = sim_lq(coal_db)

plot(simlq0, xlab = "X", ylab="Y", title="Lithotype Quality Index (Simulated)")
plot(db.locate(coal_db, "LQ", "z"), add = T)
```


# 6. Conditional simulation of the accumulation A 

Where $LQ(x) = 0$,  $\mu = 5000$ and $\sigma^{2} = 250000$

Where $LQ(x) = 1$, $\mu = 15000$ and $\sigma^{2} = 9\times10^6$.

Both $A_0$ and $A_1$ have Gaussian covariance with range 10. 

```{r warning=FALSE}

sim_a0 <- function(indb, nsim=1){
  
  indb = db.locate(indb, "Accumulation", "z")
  
  #Selecting the data for condition
  coal_a0 = db.sel(indb, LQ==1)
  gaus_a0 = model.create("Gaussian", range=10, sill=250000)
  outdb = db.create(x0=c(0,0),dx=c(1,1),nx=c(200,100))
  neighb = neigh.create(ndim = 2, type = 0)
  sim =  simtub(dbin = coal_a0,dbout=outdb,model=gaus_a0,neigh = neighb,mean=5000,seed=939452503, nbsimu = nsim)
  return(sim)
  
}


sima0 = sim_a0(coal_db)

plot(sima0, pos.legend=0, xlab="X", ylab="Y", title="Accumulation A_0 (Simulated)")
plot(db.locate(coal_db, "Accumulation", "z"), add = T)
```
```{r warning=FALSE}

sim_a1 <- function(indb, nsim=1){
  
  coal_a1 = db.locate(indb, "Accumulation", "z")
  coal_a1 = db.sel(indb, LQ==2)
  gaus_a1 = model.create("Gaussian", range=10, sill=9000000)
  outdb = db.create(x0=c(0,0),dx=c(1,1),nx=c(200,100))
  neighb = neigh.create(ndim = 2, type = 0)
  sim_a1 =  simtub(dbin = coal_a1,dbout=outdb,model=gaus_a1,neigh = neighb,mean=15000,seed=707693329, nbsimu = nsim)
  return(sim_a1)

}

sima1 = sim_a1(coal_db)

plot(sima1, xlab="X", ylab="Y", title="Accumulation A_1 (Simulated)")
plot(db.locate(coal_db, "Accumulation", "z"),add=T)
```


```{r}

#Pull out vectors of simulated 
simulated_a0 = sima0@items$Simu.Accumulation.S1
simulated_a1 = sima1@items$Simu.Accumulation.S1

#Retrieve the simulated LQ index
#Turn into binary selector/mask
mask_1 = simlq0@items$SimPGS.LQ.SFac1 - 1
#Reverse
mask_0 = 1 - mask_1

#Check - make sure the 
all(mask_0 + mask_1 == 1)

pw_map = mask_0 * simulated_a0 + mask_1 * simulated_a1


join_sims_zonal <- function(sim0, sim1, zones){
  
  mask_1 = zones - 1
  mask_0 = 1 - mask_1
  pw_map = mask_0 * sim0 + mask_1 * sim1
  return(pw_map)
  
}

simulated_a0 = sima0@items$Simu.Accumulation.S1
simulated_a1 = sima1@items$Simu.Accumulation.S1

jsim = join_sims_zonal(simulated_a0,simulated_a1,simlq@items$SimPGS.LQ.SFac1)

plot(db.locate(db.create(jsim,x0=c(0,0),dx=c(1,1),nx=c(200,100)),"jsim","z"))



```

# 7. Calorific value

Compute and plot the calorific value $C(x) = \frac{A(x)}{T(x)}$.

```{r}

cx = pw_map/simu@items$Simu.Thickness.S1

cxdb = db.create(cx,x0=c(0,0),dx=c(1,1),nx=c(200,100))
cxdb = db.locate(cxdb, 4, "z")
plot(cxdb)

```

# 8. Value of $C_{v}$ on each panel

Value of $C_{v}$ on each panel averaging over the 100 contained values. 

```{r warning=FALSE}
cgrid = db.create(x0=c(0,0),dx=c(10,10),nx=c(20,10))

plot(blockstat(cgrid, cxdb, fun='mean'))

panel_q =blockstat(cgrid, cxdb, fun='mean')
```

# 9. Quality tonnage and mean quality above cutoff curves

Quality tonnage and mean quality above cutoff for cut-off varying from 1000 to 2500. 

```{r}
hist(panel_q@items$Stats.mean)
```

Replicates the two modes, dispersion of the distribution. 


## The 

```{r}
sel = selectivity(panel_q, cuts = seq(1000,2500,50), flag.plot = F)

selectivity.curve(sel,"Z","T")

plot(sel[,1], sel[,2], type='l')
```

Interpreted in terms of the distribution, 
Obviously one iteration so it's not very 



```{r}
plot(sel[,1],sel[,5],type = 'l')
```



# 10. Simulations of Calorific Value

100 simulations

```{r warning=FALSE}
sthicks = sim_thickness(coal_db, nsim = 100)
slqs = sim_lq(coal_db, nsim = 100)
sima0 = sim_a0(coal_db, nsim = 100)
sima1 = sim_a1(coal_db, nsim = 100)
```

Calculating statistics on individual simulations

```{r warning=FALSE}
for (idx in 4:103){
  
  thickn = sthicks@items[[idx]]
  simlq = slqs@items[[idx]]
  simulated_a0 = sima0@items[[idx]]
  simulated_a1 = sima1@items[[idx]]
  ajsim = join_sims_zonal(simulated_a0,simulated_a1,simlq)
  
  cx = ajsim/thickn
  
  cxdb = db.create(cx,x0=c(0,0),dx=c(1,1),nx=c(200,100))
  cxdb = db.locate(cxdb, 4, "z")
  
  cgrid = db.create(x0=c(0,0),dx=c(10,10),nx=c(20,10))
  panel_q = blockstat(cgrid, cxdb, fun='mean')
  
  sel = selectivity(panel_q, cuts = seq(1000,2500,50),flag.plot = F)
  
  if (idx == 4){
    
    z = sel[,1]
    
    sels = vector("list", 103)
    sels[[idx]] = sel
    
  } else {
    sels[[idx]] = sel

  }
}



Mmatrix = sels[[4]][,5]
Tmatrix = sels[[4]][,2]

for (idx in 5:103){
  Mmatrix = cbind(Mmatrix, sels[[idx]][,5])
  Tmatrix = cbind(Tmatrix, sels[[idx]][,2])
}

```


# 11. Empirical 95% Confidence Interval curves


```{r}

mmu = rowMeans(Mmatrix, na.rm = T, dims = 1)
plot(z, mmu, type = 'l')

mvar = apply(Mmatrix, 1, var, na.rm = T)
msd = sqrt(mvar)

lines(z, mmu + 1.96*msd, col="red")
lines(z, mmu - 1.96*msd, col="red")
```
The standard 95% CI steady until cutoff $z$ near 2000, when the distribution of curves begins to widen. Blocks over 2000 are rare, many simulations will not produce any blocks of this value or higher. The quotient of the distributions, destroyed by averaging. In fact, the high calorific value blocks occur more in simulations where the high accumulation occurs with the low thickness, a further element of chance.


```{r}
Tm = rowMeans(Tmatrix, na.rm = T, dims = 1)
plot(z, Tm, type = 'l')



Tvar = apply(Tmatrix, 1, var, na.rm = T)
Tsd = sqrt(Tvar)

lines(z, Tm + 1.96*Tsd, col="red")
lines(z, Tm - 1.96*Tsd, col="red")
```

The tonnage cutoff (quality tonnage) curve decreases with increasing cutoff. High grade blocks. 
The curve transitions slightly between cutoff 1500 and cutoff 2000 where we pass from the area of the distribution dominated by A0 to the area dominated by A1. Drawing from two distributions to drawing from one
As seen in the mean quality above cutoff curve, not many simulations producing blocks in this range. Many simulations will simply have zero at this, driving down the variance substantially. 
The effect of drawing from a smaller set of blocks outweighs the larger variance of the higher grade blocks.
