---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---



```{r}
rm(list=ls())

coal_df = read.csv(file = 'dataset_Nicholas-Tony.csv')


coal_df$LQ = coal_df$LQ + 1

coal_df
```
```{r}
plot(coal_df$x, coal_df$y, cex = coal_df$Thickness/10.0)
```
```{r}
print(mean(coal_df$Thickness))
print(var(coal_df$Thickness))
```
```{r}
print(mean(coal_df$Accumulation))
print(var(coal_df$Accumulation))
```



```{r}
hist(coal_df$Thickness)
```
```{r}
hist(coal_df$Accumulation)
```

```{r}
library(RGeostats)



coal_db <- db.create(coal_df)

coal_db = db.locate(coal_db,3:4,"x")
coal_db = db.locate(coal_db,5,"z")

coal_db
plot(coal_db)
```

# 4. Conditional simulation of thickness T on the discretization grid

http://rgeostats.free.fr/html/simpgs.html

```{r}
Tmodel=model.create(vartype="Gaussian",range=20,sill=4,ndim=2)

outdb = db.create(x0=c(0,0),dx=c(1,1),nx=c(200,100))

neighb = neigh.create(ndim = 2, type = 0)

simu <- simtub(nbtuba = 450, dbin=coal_db,dbout=outdb,model=Tmodel,neigh = neighb,mean = c(10))

plot(simu)
plot(coal_db, add=T)
```
```{r}
hist(simu@items$Simu.Thickness.S1)
cgrid = db.create(x0=c(0,0),dx=c(10,10),nx=c(20,10))
#simu
plot(blockstat(cgrid, simu, fun='mean'))
```

# 5. Conditional simulation of LQ (lignite quality index)

Modelled as a thresholded GRF with constant proportion 0.5. 
Underlying GRF is standard ($\mu = 0$ and $\sigma^{2} = 1$) and Gaussian covariance with range 50. 

First, update the 

```{r}

props = table(coal_db[,"LQ"]) / coal_db$nsamples
props = c(0.5,0.5)
rule = rule.create(c("S", "F1", "F2"))
plot(rule, props=props, col=c("blue","yellow"), title="Lithotype Rule")
```

```{r}
coal_db = db.locate(coal_db,7,'z')
plot(coal_df$x, coal_df$y, pch = coal_df$LQ)
```

```{r}
voisin = neigh.create(type=0)
gaus = model.create("Gaussian", range=50)

sim2 = simpgs(dbout=outdb, dbin=coal_db, rule=rule,props=c(0.5, 0.5), model1=gaus,nbtuba=600, neigh = voisin)
coal_db
plot(sim2)
```
```{r}
plot(blockstat(cgrid, sim2, fun='mean'))
```



# 6. Conditional simulation of the accumulation A 

just continue with the simulation from before

Where $LQ(x) = 0$,  $\mu = 5000$ and $\sigma^{2} = 250000$

Where $LQ(x) = 1$, $\mu = 15000$ and $\sigma^{2} = 9\times10^6$.

Both have Gaussian covariance with range 10. 

***QUESTION:*** where to set the variance? we set the mean in the simtub function. 
Does this go in the "sill" argument?
```{r}
coal_a0 = db.sel(coal_db, LQ==1)
plot(coal_a0)
```

```{r}

coal_a0 = db.locate(coal_a0,6,"z")

gaus_a0 = model.create("Gaussian", range=10, sill=250000)
#Tmodel=model.create(vartype="Gaussian",range=20,ndim=2)


sim3 =  simtub(dbin = coal_a0,dbout=outdb,model=gaus_a0,neigh = neighb,mean=5000)

plot(sim3)

```
```{r}
coal_a1 = db.sel(coal_db, LQ==2)
gaus_a1 = model.create("Gaussian", range=10, sill=9000000)
#Tmodel=model.create(vartype="Gaussian",range=20,ndim=2)


sim4 =  simtub(dbin = coal_a1,dbout=outdb,model=gaus_a1,neigh = neighb,mean=15000)

plot(sim4)
plot(coal_a1,add=T)
```
***QUESTION***: 

Do we need to select then simulate? I think we need to SELECT then simulate after. 

Few possibilities:

-There is a way to use two different distributions based on the value of another variable (so one gaussian for A0, different for A1) that I am missing within the simulation code (Would this be )

-There is a built in way to draw data from two different databases dictated by a third

-There is not a built in way and I will have to do this myself. 
